An Experiment Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics

### 2020-01-10 paper/paper.tex
results

\hypersetup{
            pdftitle={Communicating human related spatial distributions: A comparison of spatial visualisations using visual inference.},
            pdfkeywords={statistics, visual inference, geospatial, population},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc header

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Communicating human related spatial distributions: A comparison of
spatial visualisations using visual inference.}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Choosing a visualisation method may seem to be a simple task. The
choropleth map display is commonly used for communicating spatial data.
The visualisation method used to present geospatial data will influence
the information derived by map users. Choosing an alternative display
will influence the communication of the distribution. The hexagon tile
map is presented as an alternative display. Visual inference is used to
measure the power of design. The choropleth is used as a comparison
point. The hexagon tile map display is an effective visualisation for
communicatin=g population related distributions. The hexagon tile map
display also allowed identification of the geographic distribution with
values monotonically decreasing from North West toward South East areas
of Australia.
\end{abstract}

@@

A geographic map base is commonly used to present geospatial statistics.
The statistics presented may also be aggreagated statistics for
geographic units, in this case the choropleth map is the common display
used. Choropleth maps are often used to present statistics regarding the
population. Creating a choropleth map involves drawing the
administrative boundaries and filling them with colour to communicate
the value of the statistic. The several sets of Australian statistical
areas presents an example of a heterogrenous distribution. The rural
communities in Australia operate on a much larger geographic space than
small inner city communities. This has the negative effect of
incorrectly showing the spatial distribution of the statistic,
especially when a spatial distribution is related to the size of the
areas, or the population density. An alternative display can be used to
effectively communicate a spatial distribution for a set of
heterogeneous areas.

@@

The Australian Cancer Atlas is an online interactive web tool created to
explore the burden of cancer on Australian communities. There are many
cancer types presented, and they can be explored on an individual or
aggregate level. The Australian communities are examined at the
Statistical Areas at Level 2 (SA2) (``Australian Statistical Geography
Standard (ASGS)'' \protect\hyperlink{ref-abs2016}{2018}) used by the
Australian Bureau of Statistics. Bayesian spatial smoothing has been
applied to incorporate the value of the statistics of the neighbouring
areas. Spatial smoothing allows for the protection privacy and gives
stability to the estimates. The statistics mapped are the diagnoses
(Standardised Incidence Rates) and excess deaths for each SA2,
communicated as the difference from the Australian average of the
statistics. The values of the statistic for each geographic area is
communicated through the colour used to fill the areas, the colours are
chosen for the diverging colour scheme. The Australian Cancer Atlas
communicates the trends in the distributions of cancer over geographic
space. It uses a choropleth map display and diverging colour scheme to
draw attention to relationships between neighbouring areas.

@@

Spatial visualisations communicate the distribution of statistics over
geographic space. The most common display for spatial data is the
choropleth map. However, the issue of using a choropleth map base
becomes obvious when considering population related distributions. Map
creators have the ability to present spatial statistics in alternative
displays that can highlight the population. This work aims to show that
a hexagon tile map display is a viable alternative to the geographic map
base for presenting population statistics. When presenting population
statistics on a geographic map base, the size of the regions can allow
errornous conclusions to be drawn about the state of the statistic over
the entire population. This occurs as large regions filled with a
consistent colour or pettern can draw the attention of map readers, and
small regions are not paid equal attention. A choropleth map is not the
only display that can be used for presenting geospatial data.
Alternative maps include various cartograms, and tesselated tile maps.
They allow other variables to be included in the display to highlight
the staistical values of various geographic areas.

A cartogram transforms the geographic map base. The transformation
begins with a choropleth map base, and the associated populations. There
are several algorithms for transforming, they all involve shifting
boundaries to increase or descrease the area of the geographic units on
the map depending on the statistics. The shifts of the boundaries result
in cartograms with accurate map area for each of the geographic units.
The transformation will keep boundaries of neighbouring areas connected,
but the result is unfamiliar shapes for each geographic unit.

Alternative displays make the trade off between familiar shapes and
representation of variables using the area of geographic units. The
non-contiguous cartgoram method keep the shapes of geographic units
intact, and change the size to communicate the values of statistics.
This method breaks the spatial relationship between areas, and
emphasises the difference between the statistic and the amount of land
for each geographic unit. The Dorling cartogram is another method that
does not retain the boundaries of geographic units. Each unit becomes
one circle on the cartogram dispaly, these can be sized according to the
value of a variable. Their neighbours can be illustrated by connecting
the boundaries of the circles. These cartogram methods all allow two
variables to be shown in the single display.

@@

Australia is an extreme case of heterogenous geographic units. To
communicate information regarding the small inner city geographic units
of Australian cities an alternative display must be employed. This will
emphasise the value of the statistic that is relating to areas with high
population levels. Figure @ref(fig:liver) shows the distribution of the
smoothed average amount of Liver cancer diagnoses for Australian Males
from XXX years. The hexagon tile map (b) shows orange hexagons for the
inner city Statistical Areas at Level 2 with higher than average levels
of diagnoses in the capital cities of Melbourne and Sydney. These inner
city areas are not visible on the choropleth map (a). There are still
many light blue areas in both displays. The northern Queensland, and
Northern Territory SA2's with higher than average rates are still
visible but utilise less map space on the hexagon tile map.

\hypertarget{methodology}{%
\subsection{Methodology}\label{methodology}}

To test the

@@

\begin{itemize}
\tightlist
\item
  Communicating data through visualisations
\item
  Effective displays for types of data
\item
  Protocol for testing the effectiveness
\end{itemize}

The same data were shown on a choropleth map, and on a hexagon tile map.
Comparing the results of participants who see the choropleth to those
who see a hexagon tile map will show that population related
distributions are spotted more frequently in a hexagon tile map display.

Classical statistical inference involves hypothesis testing, the process
of rejecting a null hypothesis in favour of an alternative. This
approach relies on data, the appropriate distributions and their
assumptions. Visual inference null hypothesis: independence in the
variables (absence of all features), alternative hypothesis:
Relationship between the variables (presence of some feature).

The lineup protocol is used for visual inference testing. 1. simulate
null plots 2. Insert data with structure into a random location 3. Ask
uninvolved person to select the most different plot 4. If location is
chosen correctly, the existence of a feature is significant at
\(\alpha = 1/N\).

\begin{quote}
``In this framework, plots take on the role of test statistics, and
human cognition the role of statistical tests.'' Buja et al.
(\protect\hyperlink{ref-SIEDAMD}{2009})
\end{quote}

The line up protocol involves placing a ``guilty'' data visualisation in
a lineup of ``innocents''. Where the guilty data set contains structure,
and the innocents are equivalent to a null data set. In a grid of
visualisations, an observer is asked to pick the display that is most
different, if they select the data set containing structure, they have
identified the guilty hidden within the group innocents. The guilty data
is identified as different from the innocent data with probability
\(1/m\), where \(m\) is the number of null plots plus 1 to account
account for the guilty data set. When the guilty data set is chosen, the
null hypothesis that it was innocent is rejected with a \(1/m\) chance
or type I error of being wrong.

The lineup protocol can be used in a variety of testing scenarios. The
choropleth map is best used for testing spatial structure in a data set.

\hypertarget{study-design}{%
\section{Study Design}\label{study-design}}

This study aims to answer several questions around the presentation of
spatial distributions:

@@

additional considerations when completing this experimental task
included exploration of the difficulty experienced by participants

The mean of the detection rate for choropleth map, denoted as \(\mu_C\),
and the hexagon tile maps, \(\mu_H\) will be contrasted. This leads to
the following one sided hypothesis for this study:

@@

The difference in the detection rates for the two displays will be
compared using the following:

\[ \hat\pi_C - \hat\pi_H \pm t_{1-\alpha,n-1}\sqrt{\hat\pi_C(1-\hat\pi_C)/n_C + \hat\pi_H(1-\hat\pi_H)/n_H} \]

The most common display for spatial cancer data is the choropleth map.
This will be the comparative visualisation for presenting the lineups
(Majumder, Hofmann, and Cook \protect\hyperlink{ref-VVSIALM}{2013}).
Most geographic distributions will have some degree of spatial
autcorrelation between neighbours. This feature will exist in all plots
in the lineup displays, the plot that contains the trend feature shown
in only one set of data will also be affected by spatial
autocorrelation. A reasonable amount of null plots \(N-1\) in the lineup
was chosen to ensure data is well hidden. For the detailed choropleth of
Australian SA2 areas, we set \(N = 12\) to not overwhelm participants. A
line up protocol was implemented to arrange 12 maps in each display.
Individual displays were created by a combination of map type, and
spatial trend model.

Recruited participants to be uninvolved judges with no prior knowledge
of the data to avoid discrimination or advantages. The online
crowdsource platform Figure-Eight was used to recruit participants.

The researchers contrasted the different plot designs, as hexagon
tilemap and geography in the lineups were created using the same data,
and same null positions within the lineup.

\hypertarget{the-variables-being-manipulated-and-measured}{%
\subsection{The variables being manipulated and
measured}\label{the-variables-being-manipulated-and-measured}}

@@

The levels of the factors measured in the experiment were: - Map type:
\emph{Choropleth, Hexagon tile} - Trend: \emph{Locations in three
population centres, Locations in multiple population centres, South-East
to North-West}

Factor combinations examined by each participant amount to 6 (2x3)
lineup displays. A participant did see the same data for both map types.
Four simulated sets of data were generated for each treatment. This will
generate 24 lineups (12 were geographic maps, and 12 were hexagon tile
maps). Participants will evaluate 12 lineups, 6 of each map type.
Appendix A shows the experimental design visually. For each of the six
geographic displays and six hexagon displays, two of each trend model
were shown to participants.

\begin{table}[]
\caption{The Experimental Design}
\label{tab:my-table}
\begin{tabular}{|l|l|l|}
\hline
Trend & Map type & Replicates \\ \hline
NW-SE & Choropleth & 2 \\ \hline
 & Hexagon tile & 2 \\ \hline
Three cities & Choropleth & 2 \\ \hline
 & Hexagon tile & 2 \\ \hline
All cities & Choropleth & 2 \\ \hline
 & Hexagon tile & 2 \\ \hline
\end{tabular}
\end{table}

@@

The underlying spatial correlation model was created to provide spatial
autocorrelation between neighbouring areas using the longitude and
latitude values for the Statistical Areas. formula = z \textasciitilde{}
1, locations = \textasciitilde{} longitude + latitude

Simulated spatially dependent data using the model on the centroids of
each area, for 12 null plots in 12 lineups.

12 sets of data were created. In these 12 sets of data, each of the 144
maps were smoothed several times to replicate the spatial
autocorrelation seen in cancer data sets presented in the Australian
Cancer Atlas.

@@

There were 95 participants involved in the study. We recruited
participants using the Figure-Eight crowd source platform by advertising
this survey to participants that fulfilled the following crtieria:

\begin{itemize}
\tightlist
\item
  level 2 or level 3 on the Figure-Eight Platform.
\item
  at least 18 years old
\end{itemize}

Participants then selected our task from the list of tasks available to
them.

Each participant was trained using three test displays orienting them to
the evaluation task. Participants then proceeded to the survey, this
involved evaluating 12 displays.

@@

@@

The survey responses from participants were kept only if the participant
submitted answers for all 12 displays. This resulted in 92 participants.

The contributors who gave various plot choices and reasons for the
twelve displays were kept. The contributors who detected no plots
correctly were analysed further. Three of these contribuotrs gave no
choices for any of the twelve displays. They were also removed for the
rest of the analysis.

Of the 92 participants, 67 were male, and 25 female. Only two of the
participants had lived in Australia before. The education level achieved
by the participants was 70 Bachelors or Masters degree.

\hypertarget{accuracy}{%
\section{Accuracy}\label{accuracy}}

The detection rate for participants reporting the real data trend model
is used to calculate the accuracy of the responses of participants.

@@

The detection rates are usually higher when participants viewed the
hexagon tile map displays. The difference was greater when viewing a
three cities trend model. When the trend model affected all cities
participants also detected the data display in the lineup of null plots
more often. The North West to South East data plot was detected with
similar rates when viewing the choropleth and hexagon tile map displays,
though for three of the replicates the participants detected the data
display more often.

A t-test shows the difference between the detection rates for the two
types of displays. The value of 0.0041593 shows that it is very unlikely
the difference is due to chance.

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}

The proportion of correct and incorrect responses within each 5 second
time window. For the NW-SE distribution, it is equally likely that quick
responders will be correct or incorrect. It is more likely that
participants picked the correct response if they took longer to respond.
For the Choropelth

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/reason-1.pdf}
\caption{The amount of participants that selected each reason for their
choice of plot when looking at each trend model shown in Choropleth and
Hexagon Tile maps. The facets show whether or not the choice was
correct.}
\end{figure}

@@

@@

The choices made by participants are examined in Figure
@ref(fig:choices). Participants were misled by the choropleth display,
but not the hexagon display for all cities displays except (2). The maps
with a North West to South East trend was chosen with much greater
frequency in all displays. All of three cities displays, except (4),
were detected in the hexagon display. All except one lineup had at least
one participant select the correct map in the lineup as shown in Figure
@ref(fig:choices\}.

@@

\begin{tabular}{l|r|r|r|r|l|l}
\hline
\_ & terms & Estimate & Error & Stat & p.value & sig\\
\hline
Design & Intercept & 0.0217400 & 0.1474507 & 0.1474391 & 0.8828 & \\
\hline
 & Hex & 0.4200928 & 0.2110904 & 1.9901083 & 0.0466 & *\\
\hline
Trend & Three cities & -3.2519793 & 0.4124809 & -7.8839514 & 0.0000 & ***\\
\hline
 & All cities & -1.2398974 & 0.2293327 & -5.4065450 & 0.0000 & ***\\
\hline
Interactions & Hex : Three cities & 2.3683138 & 0.4645163 & 5.0984512 & 0.0000 & ***\\
\hline
 & Hex : All cities & 1.0825772 & 0.3123987 & 3.4653698 & 0.0005 & ***\\
\hline
\end{tabular}

@@

The estimated marginal mean is the fitted value from a model over the
treatment effects. The treatment effects considered in this model are
map type displayed, trend model included, and the replicate. The
\texttt{emmeans} package is used to calculate the estimated marginal
mean. The value, between 0 and 1, can be interpretted as the probability
of detecting the real data plot in the lineup, for each combination of
map type, trend model and replicate.

\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}

\begin{verbatim}
## # A tibble: 1 x 6
##   sigma logLik   AIC   BIC deviance df.residual
##   <dbl>  <dbl> <dbl> <dbl>    <dbl>       <int>
## 1 0.430  -675. 1367. 1407.    1321.        1096
\end{verbatim}

\begin{verbatim}
## # A tibble: 8 x 5
##   term                       estimate std.error statistic group      
##   <chr>                         <dbl>     <dbl>     <dbl> <chr>      
## 1 (Intercept)                   0.505    0.0340     14.9  fixed      
## 2 typeHex.                      0.103    0.0448      2.31 fixed      
## 3 trendThree Cities            -0.467    0.0448    -10.4  fixed      
## 4 trendAll Cities              -0.277    0.0448     -6.19 fixed      
## 5 typeHex.:trendThree Cities    0.250    0.0633      3.95 fixed      
## 6 typeHex.:trendAll Cities      0.239    0.0633      3.77 fixed      
## 7 sd_(Intercept).contributor    0.118   NA          NA    contributor
## 8 sd_Observation.Residual       0.430   NA          NA    Residual
\end{verbatim}

>>>>

\hypersetup{
            pdftitle={An Experiment Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics},
            pdfkeywords={statistics, visual inference, geospatial, population},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc header
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{An Experiment Comparing the Effectiveness of the Choropleth Map with a
Hexagon Tile Map for Communicating Cancer Statistics}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Choosing a visualisation method may seem to be a simple task. The
choropleth map display is commonly used for communicating spatial data.
However, when used to communicate the spatial distribution of a
statistic the sizes of areas can lead to midinterpretation of the
distribution. The visualisation method used to present geospatial data
will influence the information derived by map users. Choosing an
alternative display will influence the communication of the
distribution. The hexagon tile map is presented as an alternative
display. Visual inference is used to measure the power of design. The
choropleth is used as a comparison point. The hexagon tile map display
is an effective visualisation for communicatin=g population related
distributions. The hexagon tile map display also allowed identification
of the geographic distribution with values monotonically decreasing from
North West toward South East areas of Australia.
\end{abstract}

@@

This study contrasts the use of two displays for geospatial statistics,
the traditional choropleth map and an alternative tessellated hexagon
tile map. The choropleth map is the common choice for visualising
aggregated statistics and is familiar because of common administrative
boundaries. A hexagon tile map forgoes the familiar boundaries, instead
representing each geographic unit as an equally sized hexagon. This
alternative display is suggested for communicating distributions that
relate to the population values or density of the population in each
geographic unit. This study uses the lineup protocol for visual
inference testing of the recognition of spatial distributions. The
Australian Cancer Atlas explores spatial distributions of cancer
incidence and the burden on Australian communities. The spatial
distributions of the cancers explored in this atlas inspired the trend
models presented to participants in this study.

@@

The Australian Cancer Atlas is an online interactive web tool created to
explore the burden of cancer on Australian communities. There are many
cancer types to be explored individually or aggregated. The Atlas
examines Australian communities at the Statistical Areas at Level 2
(SA2) (``Australian Statistical Geography Standard (ASGS)''
\protect\hyperlink{ref-abs2016}{2018}) used by the Australian Bureau of
Statistics. Bayesian spatial smoothing was applied to integrate the
values of neighbouring areas. Spatial smoothing protects the privacy and
gives stability to the estimated amounts in each SA2. Users of the Atlas
can explore the difference of the statistic of each SA2 from the
Australian average, for both diagnoses (Standardised Incidence Rates)
and excess deaths. It communicates the value of the statistic for each
SA2 geographic unit through colour, by filling the areas using the
diverging colour scheme ranging from dark blue to red. The Australian
Cancer Atlas allows users to explore the patterns in the distributions
of cancer statistics over the geographic space of Australia. It uses a
choropleth map display and diverging colour scheme to draw attention to
relationships between neighbouring areas.

@@

Spatial visualisations communicate the distribution of statistics over
geographic space. The most common display for spatial data is the
choropleth map. The choropleth map is the common display used to present
aggregated statistics for geographic units such as the population.
Creating a choropleth map involves drawing the administrative boundaries
and filling them with colour to communicate the value of the statistic.
However, using a choropleth map base introduces biases when considering
population related distributions. Many sets of Australian statistical
areas have a large disparity between the smallest and largest units. The
rural communities in Australia operate on a much larger geographic space
than small inner-city communities. When a spatial distribution is
related to the size of the areas, or the population density, the size of
the regions can allow erroneous conclusions to be drawn about the state
of the statistic over the entire population, when presenting the
statistics on a geographic map base. This occurs as large regions filled
with a consistent colour or pattern can draw the attention of map
readers, and small regions are not paid equal attention. An alternative
display can be used to effectively communicate a spatial distribution
for a set of heterogeneous areas. This work aims to show that a hexagon
tile map display is a viable alternative to the geographic map base for
presenting population statistics.

A choropleth map is not the only display that can be used for presenting
geospatial data. Alternative maps include various cartograms and
tessellated tile maps. They allow other variables to be incorporated in
the display to highlight the statistical values of various geographic
areas.

A cartogram transforms the geographic map base. The transformation
begins with a choropleth map, and the values of a statistic such as
population, for each geographic unit. There are several algorithms for
transformation; they involve shifting the boundaries of geographic
units, using the value of the statistic to increase or decrease the area
taken by the geographic unit on the map. The changes to the boundaries
result in cartograms that accurately communicate population by map area
for each of the geographic units. The transformation will keep
boundaries of neighbouring areas connected, but the result is unfamiliar
shapes for each geographic unit.

Alternative displays make the trade off between familiar shapes and
representation of variables using the area of geographic units. The
non-contiguous cartgoram method keeps the shapes of geographic units
intact, and change the size of the shape to communicate the values of
statistics. This method disconnects areas when shrinking boundaries, and
breaks the spatial relationship between areas. However, the empty space
created emphasises the comparative density of the statistic, the
difference between the statistic and the amount of land for each
geographic unit. The Dorling cartogram is another method that does not
retain the boundaries of geographic units. Each unit becomes one circle
on the cartogram display; the circles are sized according to the value
of a the statistic, such as population. The neighbour relationships can
be illustrated by connecting the boundaries of the circles. These
cartogram methods all allow two variables to be shown in the single
display.

@@

Australia is an extreme case of heterogenous geographic units, with
large difference between the smallest and largest geographic units
across various granularities. To communicate information regarding the
small inner-city geographic units of Australian cities an alternative
display must be employed. This will emphasise the value of the statistic
for areas with high population. Figure @ref(fig:liver) shows the
distribution of the smoothed average amount of Liver cancer diagnoses
for Australian Males from 2005-2014. These inner-city areas are not
visible on the choropleth map (a). The hexagon tile map (b) shows orange
hexagons for the inner-city SA2s with higher than average levels of
diagnoses in the capital cities of Melbourne and Sydney. The northern
Queensland, and Northern Territory SA2's with higher than average rates
are still visible but use less map space on the hexagon tile map. There
are still many light blue areas in both displays.

@@

The hexagon tile map is suggested as an alternative display. To test the
effectiveness of these displays there is a formal testing framework.
Visual inference considers the communication of data or data summaries
through visualisations. It considers these visualisations to be visual
statistics, with observable features.

Classical statistical inference involves hypothesis testing, the process
of rejecting a null hypothesis in favour of an alternative. This
approach requires appropriate data, with the appropriate distributions
and their assumptions. The lineup protocol was formalised as a test for
the effectiveness of a visualisation. Rather than use classical
statistical tests, human participants are tasked with finding the
display that shows structure between the variables. If the structure is
not detected, then it is not significantly different from the null data.

Visual inference uses the general hypothesis structure:

Null hypothesis: There is NO relationship between the variables,
Alternative hypothesis: There is a relationship between the variables.

To test these hypotheses, the line up protocol randomly places a
``guilty'' data visualisation in a lineup of ``innocents''. Where the
guilty data set contains structure, and the innocents data sets do not.
In a grid of visualisations, an observer is asked to pick the display
that is most different, if they select the data set containing
structure, they have identified the guilty hidden within the innocents
(Hofmann et al. \protect\hyperlink{ref-GTPCCD}{2012}). The guilty data
is identified as different from the innocent data with probability
\(1/m\), where \(m\) is the number of null plots plus 1 to account
account for the guilty data set. When the guilty data set is chosen, the
null hypothesis that it was innocent is rejected with a \(1/m\) chance
or type I error of being wrong.

The lineup protocol can be used in a variety of testing scenarios for
visualisations. Wickham et al. (\protect\hyperlink{ref-GIIV}{2010})
suggest the choropleth map for testing spatial structure in a data set.
The lineup protocol allows for flexibility in the definition of null
data. In the case of spatial visualisations it is likely that neighbours
are related to some degree. To account for this, null data sets can be
generated by sampling from a known model that accounts for spatial
covariance.

To contrast the effectiveness of two displays, the data set of null data
plots with a hidden real data plot can be produced using the different
plot designs. The time taken by participants to evaluate the same data
in the different displays can be compared. Their accuracy in detecting
the real data plot can also be calculated and contrasted.

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

This study aimed to answer two key questions around the presentation of
spatial distributions:

@@

Additional considerations when completing this experimental task
included exploration of the difficulty experienced by participants, and
the certainty they felt about their decision.

The mean of the detection rate for choropleth map, denoted as \(\mu_C\),
and the hexagon tile maps, \(\mu_H\) will be contrasted. This leads to
the following one sided hypothesis:

@@

The difference in the detection rates for the two displays will be
compared using a two sample unpaired t-test.

Participants were recruited for a survey style task to test the
effectiveness of the hexagon tile map display. It was expected that the
participants were uninvolved judges that had no prior knowledge of the
data, to avoid discrimination or advantages. The Figure-Eight crowd
source platform was used to advertise the survey to potential
participants that had acheived level 2 or level 3 on the Figure-Eight
Platform. The participants were chosen as they have experience with
survey tasks that involve evaluating images. They participants on the
platform are also required to be at least 18 years old.

Participants were able to choose to participate by selecting this task
from the list of tasks available to them. There were 95 participants
involved in the study. All of these participants read the introductory
materials and viewed example questions before proceeding to the survey.
Each participant was trained using three test displays orienting them to
the evaluation task, these displays are shown in supplementary materials
section regarding \protect\hyperlink{training}{training}. All
participants who volunteered to take part were compensated for their
time via the payment system of Figure-Eight.

Participants were allocated to either group A or group B when they
proceeded to the survey web application, hosted externally from the
Figure-Eight website.

\hypertarget{variables}{%
\subsection{Variables}\label{variables}}

@@

The levels of the factors measured in the experiment were:

\begin{itemize}
\tightlist
\item
  Map type: \emph{Choropleth, Hexagon tile}
\item
  Trend: \emph{Locations in three population centres, Locations in
  multiple population centres, South-East to North-West}
\end{itemize}

Each Group did not see the same data for both map types. Four simulated
sets of data were generated for each treatment. This will generate
twenty-four lineups (twelve lineups of geographic maps, and twelve
lineups of hexagon tile maps). Participants evaluated twelve lineups,
siz of each map type. For each of the six geographic displays and six
hexagon displays, two of each trend model were shown to participants.

\begin{table}

\caption{\label{tab:exp_design}Experimental design.}
\centering
\begin{tabular}[t]{llll}
\toprule
Group & Trend & Choro. & Hex.\\
\midrule
A & NW-SE & 1, 2 & 3, 4\\
A & Three Cities & 1, 2 & 3, 4\\
A & All Cities & 2, 4 & 1, 3\\
B & NW-SE & 3, 4 & 1, 2\\
B & Three Cities & 3, 4 & 1, 2\\
\addlinespace
B & All Cities & 1, 3 & 2, 4\\
\bottomrule
\end{tabular}
\end{table}

@@

The lineup protocol allows a known model to be imposed on the null data
set plots. The underlying spatial correlation model was created to
provide spatial autocorrelation between neighbouring areas using the
longitude and latitude values for the Statistical Areas.

\[z = 1\] \[locations = longitude + latitude\]

The null model imposed suggests that neighbours are related. The
randomness imposed was smoothed to mirror the practice taken by the
Australian Cancer Atlas. This smoothing allowed neighbours to be related
to each other and show distributions similar to the Liver cancer
distribution shown in @ref(fig:liver).

Spatially dependent data sets were simulated using the variogram model
on the centroids of each geographic unit. Twelve simulations from the
variogram model were created for each of the twelve lineups.

In these 12 sets of data, each of the 144 maps were smoothed several
times to replicate the spatial autocorrelation seen in cancer data sets
presented in the Australian Cancer Atlas.

@@

Replicate 3 of the All Cities distribution is shown in
@ref(fig:compare). The same data set is shown in both displays, the
difference is the land area of each SA3 is coloured in the choropleth
and the hexagon representing the SA3 is coloured in the Hexagon Tile
Map. Many maps in the Choropleth lineup (a) are coloured yellow, two are
half light blue, and two are obviously blue. Only one of the maps looks
red.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/compare-1.pdf}
\caption{Comparing the same data presented in two map displays. The
choropleth map display in (a)}
\end{figure}

@@

The choropleth map was used as the comparative visualisation for
presenting the lineups (Majumder, Hofmann, and Cook
\protect\hyperlink{ref-VVSIALM}{2013}) as this is the common display for
spatial cancer data. Geographic distributions usually have some degree
of spatial autcorrelation between neighbours. This feature was
incorporated in all maps shown in the lineup displays, the map that
contained the trend feature shown in only one set of data was also
affected by spatial autocorrelation. A line up protocol was implemented
to arrange \(N\) maps in each display. A reasonable amount of null plots
\(N-1\) in the lineup was chosen to ensure the real data map was well
hidden. A reasonable number of plots to show in each lineup, \(N = 12\)
was chosen to not overwhelm participants due to the detailed choropleth
maps of Australian SA3 areas.

The same data was visualised on a choropleth map, and on a hexagon tile
map, however in this study the participants only saw one of these two
displays. The accuracy and times taken were aggregated for Groups and.
Comparing the results of participants who see the choropleth to those
who see a hexagon tile map will show that population related
distributions are spotted more frequently in a hexagon tile map display.

@@

Responses from 95 participants were collected. Three participants
provided no answers for any task, and their data was removed. Set A was
evaluated by 42 participants, and 53 evaluated set B. This resulted in
1104 evaluations, corresponding to 92 subjects, each evaluating 12
lineups, that are analysed on accuracy and speed.

\hypertarget{participant-demographics}{%
\subsection{Participant demographics}\label{participant-demographics}}

Of the 92 participants, 67 were male, and 25 female. Most participants
(56) had a Bacherlors degree, 13 had a Masters degree, and the remaining
23 had high school diplomas.

\hypertarget{accuracy}{%
\subsection{Accuracy}\label{accuracy}}

Figure @ref\{detection\_compare\} display the average detection rates by
type of plot, separately for each trend model. Evaluations on the same
data set are connected by a line segment. For both the Three Cities and
All Cities trend models performance was substantially better for the
hexagon tile map. Viewers detected the data plot substantially more
often with the hexagon tile map. One replicate for the All Cities group,
had similar detection rates for both plot types, and this is discussed
in a later section. Surprisingly, participants could also detect the
gradual spatial trend in the NW-SE group from the hexagon tile map. We
expected that the choropleth map would be superior for the type of
spatial pattern, but the data suggests the hexagon tile map performs
equally as well.

Table \ref{tab:desc_stats} shows the means and standard deviations for
type of plot and each trend model.

@@

\begin{table}

\caption{\label{tab:desc_stats}}
\centering
\begin{tabular}[t]{lccc}
\toprule
Type & NW-SE & Three Cities & All Cities\\
\midrule
Choro. & 0.505 & 0.038 & 0.228\\
 & (0.5) & (0.19) & (0.42)\\
Hex. & 0.609 & 0.391 & 0.571\\
 & (0.49) & (0.49) & (0.5)\\
\bottomrule
\end{tabular}
\end{table}

Table XXX is a summary of the mixed effects model. Blah, blah, what we
learn.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/beeswarm-1.pdf}
\caption{The beeswarm plots show the distribution of the time taken to
submit a response for each combination of trend and type of display. The
time taken shown by the height of the point for each evaluationq1. The
height of the histogram bars show how many evaulations were submitted
within each time window. The orange regions of the bars show the amount
of correct detections, and the green regions are the amount of incorrect
detections.}
\end{figure}

The proportion of correct and incorrect responsesis shown by the width
used to distribute the points horizontally. Each point is randomly
placed, but regions with more points require points to be placed further
from center. For the NW-SE distribution, it is equally likely that quick
responders will be correct or incorrect. It is more likely that
participants picked the correct response if they took longer to respond.
For the Choropelth

@@

\begin{table}[H]
\centering
\begin{tabular}{llll}
\toprule
Trend & Detected & Choro. & Hex.\\
\midrule
NW-SE & No & trend & clusters\\
NW-SE & Yes & trend & clusters\\
Three Cities & No & trend & clusters\\
Three Cities & Yes & consistent & clusters\\
All Cities & No & trend & clusters\\
\addlinespace
All Cities & Yes & clusters, consistent & clusters\\
\bottomrule
\end{tabular}
\end{table}

@@

The choices made by participants are examined in Figure
@ref(fig:choices). Participants were misled by the choropleth display,
but not the hexagon display for all cities displays except (2). The maps
with a North West to South East trend was chosen with much greater
frequency in all displays. All of three cities displays, except (4),
were detected in the hexagon display. All except one lineup had at least
one participant select the correct map in the lineup as shown in Figure
@ref(fig:choices).

@@

One lineup had extremely similiar detection rates for the Choropleth and
Hexagon displays.

\begin{figure}
\centering
\includegraphics{figures/aus_cities_3_geo.png}
\caption{``The lineup of choropleth map displays shows a distribution
that affects all capital cities. The values for the inner city areas in
the capital cities result in them being coloured red. However, the
colour blue dominates the display as the large rural areas are filled.
Replicate (2) for all cities.''}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/aus_nwse_6_hex.png}
\caption{``A North-West to South East trend was hidden in this lineup of
hexagon tile map displays.''}
\end{figure}

@@

\begin{tabular}{rcrrr}
\toprule
Term & Estimate & Error & Stat & p.value\\
\midrule
Intercept & 0.02 & 0.15 & 0.15 & 0.88\\
Hex & 0.42$^{*  }$ & 0.21 & 1.99 & 0.05\\
Three cities & -3.25$^{***}$ & 0.41 & -7.88 & 0.00\\
All cities & -1.24$^{***}$ & 0.23 & -5.41 & 0.00\\
Hex : Three cities & 2.37$^{***}$ & 0.46 & 5.10 & 0.00\\
\addlinespace
Hex : All cities & 1.08$^{***}$ & 0.31 & 3.47 & 0.00\\
\bottomrule
\end{tabular}

@@

\includegraphics{paper_files/figure-latex/glm2-1.pdf}

\% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard
University. E-mail: hlavac at fas.harvard.edu \% Date and time: Fri, Jan
10, 2020 - 11:55:55
